---
title: "Geo Experiment"
author: "Chris Comiskey"
date: "6/17/2019"
output: html_document
---

```{r, message=FALSE, echo = FALSE, warning=FALSE}
library(readr)
library(devtools)
library(GeoexperimentsResearch)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)

nl_geo <- readr::read_csv(
  "nl_geo.csv", col_types = cols(X1 = col_skip(), date = col_date(format = "%Y-%m-%d"))
  )
```


```{r, echo = FALSE}
geo_ts <- GeoTimeseries(nl_geo, metrics = c('Cost', 'Revenue'))

phases <- ExperimentPeriods(c("2019-01-01", "2019-03-06", "2019-05-19"))

geo_assign <- nl_geo %>%
  select(geo, geo.group) %>%
  distinct() %>%
  GeoAssignment()

geo_obj <- GeoExperimentData(geo.timeseries = geo_ts, 
                             periods = phases, 
                             geo.assignment = geo_assign)

```

## NL Revenue Results

Time based regression yields these results: estimated revenue increase, confidence interval lower and upper bounds, standard error (estimated standard deviation), level of confidence interval.

```{r, echo=FALSE, warning=FALSE}
tbr_rev <- DoTBRAnalysis(geo_obj, 
                         response = 'Revenue',
                         model = 'tbr1',
                         pretest.period = 0,
                         intervention.period = 1,
                         control.group = 1,
                         treatment.group = 2)

tbr_rev_summary = summary(tbr_rev, level = 0.95, interval.type = "two-sided")
tbr_rev_summary[-c(2, 7:9)]
```

The esimated revenue increase due to the treatment is €49470, with 95% confidence interval (€18178, €80762).

Time based regression results:

```{r , warning = FALSE, echo = FALSE}
plot(tbr_rev) 
```

* The top panel shows the *observed* treatment time series (red) and the estimated counterfactual. 
* The middle panel shows the pointwise, daily difference between the observed and estimated counterfactual. 

* The bottom panel shows the sum of these differences. 

The shaded region around the counterfactual line, in each panel, depicts the counterfactual 95% confidence interval

## NL iROAS Results

