---
title: "Calculating Indexes with cpir"
author: "Chris Comiskey, Open Data Group"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating Indices with cpir}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
library(cpir)
```

This vignette shows how to use two functions, using three included datasets, and highlights our approach to organizing index and subindex dates.

## Functions

**cpir** contains two functions: 

- `weightr` for calculating relative important weights (RIWs); 
- `indexr` for calculating aggregate indices, such as the titular aggregate index  CPI (Consumer Price Index). 

Implicit in these functions is an approach to arranging the index and subindex dates. 

## Data

**cpir** includes three datasets, which this vignette uses to show how `weightr` and `indexr` functionality:

1. `index` an example aggregate index; actual CPI, by date
2. `subindices` an example set of subindices; actual expenditure class subindices, by date
3. `weights` example RIWs; actual December 2009 RIWs

Let's take a look at a bit of each dataset.

#### 1. `index`
This dataset is a 100 $\times$ 2 data frame, where each row consists of a date and corresponding index value. Here are the first six rows:
```{r, echo = FALSE}
head(index)
```

#### 2. `subindices`
This dataset is a 100 $\times$ 69 data frame, where each row consists of a date and 68 subindex values. The subindex column names are the official [BLS](https://www.bls.gov/help/hlpforma.htm) "item codes".  Here are the first five rows and five columns:
```{r, echo=FALSE}
subindices[1:5, 1:6]
```

#### 3. `weights`
This dataset is a 1 $\times$ 69 data frame, where column one is the date of the category weights occupying the remaining 68 columns. Here are the first five rows and five columns:
```{r, echo=FALSE}
weights[1,1:6]
```
Note that the order of the BLS item codes matches in `subindices` and `weights`.

## RIWs with `weightr`
The function `weightr` applies a [BLS](https://www.bls.gov/opub/hom/pdf/homch17.pdf) formula to calculate the month $t$ weight of a single subindex. Given weights $\text{RIW}_{t-1}$; subindices $\text{EC}_{t}$, $\text{EC}_{t-1}$; and aggregate indices $\text{CPI}_{t}$, $\text{CPI}_{t-1}$; the formula for the month $t$ weight is:

$$ \text{RIW}_{t} = \text{RIW}_{t-1} \left(
  \frac{\text{EC}_{t}}{\text{EC}_{t-1}}
  \right) \left( \frac{\text{CPI}_{t-1}}
  {\text{CPI}_{t}} \right).$$
  
Accordingly, `weightr(...)` takes the following five arguments.

Argument         | Variable             | Description
-----------------|----------------------|-------------
`w0`             | $\text{RIW}_{i,t-1}$ | Reference weight
`sub0`           | $\text{EC}_{i,t-1}$  | Reference subindex
`index0`         | $\text{CPI}_{t-1}$   | Reference aggregate index
`sub1`           | $\text{EC}_{i,t}$    | Current subindex
`index1`         | $\text{CPI}_{t}$     | Current aggregate index
.
.
****** **REITERATE THAT ITEM_CODE ORDER MATCHES** *****
.
.

We can calculate the second month ("2010-01-01") weights.
```{r, eval=FALSE}
weightr(
  w0 = weights[1,-1],      # reference month weights vector
  sub0 = subindices[1,-1], # reference month subindices vector
  index0 = index[1,2],     # reference month index
  sub1 = subindices[2,-1], # current month subindices vector
  index1 = index[2,2]      # current month index
  )
```
Notice that all reference dates match, for reference and current months respectively; i.e.
```{r}
weights[1,'Date'] == subindices[1,'date'] &
subindices[1,'date'] == index[1,'date'] &
subindices[2,'date'] == index[2,'date']
```
That is, the reference weights, subindices, and aggregate index all come from the same point in time; and the current subindices come from the same point in time as the current index.

## Aggregate Index with `indexr`
The function `indexr` applies a [BLS](https://www.bls.gov/opub/hom/pdf/homch17.pdf) formula to calculate a month $t$ aggregate index, such as CPI. Let $t$ represent time, and $i$ index subindices. Then given weights $\text{RIW}_{i,t-1}$; subindices $\text{EC}_{i,t}$, $\text{EC}_{i,t-1}$; and the previous aggregate indices $\text{CPI}_{t-1}$; the formula for the month $t$ aggregate index $\text{CPI}_{t}$ is:

$$ \text{CPI}_{t} = \text{CPI}_{t-1} \left(
  \sum_{i} \left[
    \text{RIW}_{i,t-1} \left(
      \frac{\text{EC}_{i,t}}{\text{EC}_{i,t-1}}
    \right)
  \right]
\right) $$

Accordingly, `indexr(...)` takes the following arguments.

Argument         | Variable             | Description
-----------------|----------------------|-------------
`w0`             | $\text{RIW}_{i,t-1}$ | Reference weights
`sub0`           | $\text{EC}_{i,t-1}$  | Reference subindices
`index0`         | $\text{CPI}_{t-1}$   | Reference aggregate index
`sub1`           | $\text{EC}_{i,t}$    | Current subindices

.
.
******* **REITERATE THAT ITEM_CODE ORDER MATCHES** *******
.
.

We can caluclate a current month's aggregate index, in this case month 2 "2010-01-01".
```{r}
indexr(
  w0 = weights[1,-1],      # reference month weights vector
  sub0 = subindices[1,-1], # reference month subindices vector
  index0 = index[1,2],     # reference month index
  sub1 = subindices[2,-1]  # current month subindices vector
  )
```
Again, it is important to note that the dates match as they should, and need to.
```{r}
weights[1,'Date'] == subindices[1,'date'] &
subindices[1,'date'] == index[1,'date']
```
That is, the reference weights, subindices, and aggregate index all come from the same point in time.
