\documentclass{article}

\title{Macroeconomic Models: \\ Weekly Update}
\author{Chris Comiskey, Open Data Group}
\date{\today}

\usepackage{natbib}
\bibliographystyle{unsrtnat}

\usepackage{fullpage}
\usepackage{ulem}

\usepackage{amsmath, amsthm, amssymb, amsfonts}
\usepackage{mathtools}
\usepackage{float}
\usepackage{bbm}

\usepackage{listings}


\begin{document}

\maketitle{}

\begin{itemize}
\item Discussing things this week, it was noted that the BLS does not provide RIWs for ELIs; for aggregating to strata indices.
\item With the previous bullet in mind, recall Motor Fuel, the Expenditure Class which we saw last week ``explains'' the most CPI month-to-month variation \footnote{according to $R^{2}$ of a simple linear regression of every EC individually}. Using this EC for an example.
\item Consider regressing the Gasoline strata\footnote{The month-to-month changes in the Gasoline Strata.} (subindex of Motor Fuel EC) on its three ELIs (i) Regular Unleaded Gasoline; (ii) Mid-grade Unleaded Gasoline; and (iii) Premium Unleaded Gasoline.
$$ \text{Gas}_{i} = \beta_{0} + \beta_{1}\text{Regular}_{i} + \beta_{2}\text{Mid}_{i} + \beta_{1}\text{Premium}_{i} + \epsilon_{i} $$
This regression gives helpful results.
\begin{verbatim}
Call:
lm(formula = diff(gas) ~ diff(reg) + diff(mid) + diff(prem), 
    data = g)
Coefficients:
             Estimate Std. Error t value Pr(>|t|)    
(Intercept) -0.004356   0.006727  -0.648    0.518    
diff(reg)    0.748003   0.009953  75.154   <2e-16 ***
diff(mid)    0.089597   0.008671  10.333   <2e-16 ***
diff(prem)   0.166801   0.009188  18.155   <2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Multiple R-squared:  0.9999,	Adjusted R-squared:  0.9999 
F-statistic: 1.148e+06 on 3 and 289 DF,  p-value: < 2.2e-16
\end{verbatim}
\item Two points worth mentioning:
  \begin{enumerate}
  \item $\text{R}^{2} \approx 1$, indicating that a linear combination of the ELIs explains the variation in the strata index. 
  \item $\sum_{i} \beta_{i} \approx 1$, as RIWs would sum to one. 
  \end{enumerate}
Together these points suggest such a regression provides estimates of the missing RIWs for aggregating ELIs to strata level indices.
\end{itemize}
\end{document}